---
title: "Wellspring Data Analysis Project Proposal"
author: "Jingtian Hu - (Instructors: Nathalie Moon and Morris Greenberg)"
format: pdf
editor: visual
---model <- rpart(activeness ~ age_years + gender + program_interest + mailing_city + parent_of_a_child_under_18 + types_of_cancer + i_have_the_following_impairment + how_have_you_heard_of_wellspring, data=added_frequency_category)
---

```{r}
library(tidyverse)
library(clock)
library(rpart)
library(partykit)
setwd("~/sta130_project/")
data <- read_csv("v1_wrangled_Member_Background.csv")
```

```{r}
added_dates_data <- data |>
     mutate(
         starting_date = date_build(member_start_year, member_start_month, ),
         ending_date = date_build(2025, 2, 1),
     )

added_membership_duration <- added_dates_data |>
  mutate(mem_duration_month = (ending_date - starting_date) |> as.numeric() / 30)

added_delivery_frequency <- added_membership_duration |> 
  mutate(present_service_deliveries_per_month = number_of_present_service_deliveries / mem_duration_month)

print(median(added_delivery_frequency$present_service_deliveries_per_month))
print(mean(added_delivery_frequency$present_service_deliveries_per_month))

added_frequency_category <- added_delivery_frequency |>
  mutate(activeness = case_when(
    present_service_deliveries_per_month <= 0.11 ~ "low",
    present_service_deliveries_per_month > 0.11 ~ "high",
  ))


exclude_na <- added_frequency_category |> filter(!is.na(i_have_the_following_impairment))

process_na <- added_frequency_category |> replace_na(list(program_interest="Unknown"))
```

```{r}
# traning without crossvalidate
model <- rpart(activeness ~ age_years + i_identify_as_lgbtq + parent_of_a_child_under_18 + i_identify_as_poc + i_have_the_following_impairment + , data=exclude_na)
predictions <- predict(model, newdata=added_frequency_category, type="class")
table(predictions, added_frequency_category$activeness)
```

```{r}
# traning with crossvalidate
model <- rpart(present_service_deliveries_per_month ~ (age_years + gender + types_of_cancer), data=final_data)

```
