---
title: "Wellspring Data Analysis Project Proposal"
author: "Jingtian Hu - (Instructors: Nathalie Moon and Morris Greenberg)"
format: pdf
editor: visual
---model <- rpart(activeness ~ age_years + gender + program_interest + mailing_city + parent_of_a_child_under_18 + types_of_cancer + i_have_the_following_impairment + how_have_you_heard_of_wellspring, data=added_frequency_category)
---

```{r}
library(tidyverse)
library(clock)
library(rpart)
library(partykit)
setwd("~/sta130_project/")
data <- read_csv("v3_wrangled_Member_Background.csv")
```

```{r}
added_dates_data <- data |>
     mutate(
         starting_date = date_build(member_start_year, member_start_month, ),
         ending_date = date_build(2025, 2, 1),
     )

added_membership_duration <- added_dates_data |>
  mutate(mem_duration_month = (ending_date - starting_date) |> as.numeric() / 30)

added_attendance_rate <- added_membership_duration |> 
  filter(number_of_present_service_deliveries + number_of_absent_service_deliveries > 0) |>
  mutate(attendance_rate = number_of_present_service_deliveries / (number_of_present_service_deliveries + number_of_absent_service_deliveries))

print(median(added_attendance_rate$attendance_rate))
print(mean(added_attendance_rate$attendance_rate))
```

```{r}
added_attendance_rate |> ggplot(aes(x=age_years, y=attendance_rate, color = gender)) + geom_point() + labs(x="ages in years", y="attendance_rate")
```


```{r}
# traning without crossvalidate
model <- lm(attendance_rate ~ age_years * gender * i_identify_as_lgbtq * parent_of_a_child_under_18 * i_identify_as_poc * i_have_the_following_impairment * how_have_you_heard_of_wellspring * types_of_cancer + program_interest + program_language_preference, data=added_attendance_rate)
summary(model)
```

```{r}
# traning with crossvalidate
model <- rpart(present_service_deliveries_per_month ~ (age_years + gender + types_of_cancer), data=final_data)
```
