---
title: "Wellspring Data Analysis Project Proposal"
author: "Jingtian Hu - (Instructors: Nathalie Moon and Morris Greenberg)"
format: pdf
editor: visual
---model <- rpart(activeness ~ age_years + gender + program_interest + mailing_city + parent_of_a_child_under_18 + types_of_cancer + i_have_the_following_impairment + how_have_you_heard_of_wellspring, data=added_frequency_category)
---

# Research problem: can we predict people's fav program type according to self-reported demographic info?

## Research method, Classification Decision Tree with rpart

## relevance:
Wellspring manager can benefit from knowing members' program preferences by
1. recommending programs to members that they are more likely to be interested in
2. tailoring particular program types to approach certain groups of members.

## Processing:

1. In service table, I transformed every service's name to exclude the subtitle (anything after ":" if any) so that classifying can be done to fewer session names.
2. Then, I grouped services by member_id and summarize that which session name was each one's favorite (judging by number of registration not succ attendances).
  - Intermediate step with GenAI, I used Google Gemini to generate a mapping from session name to program type (I checked the classification, it made sense but might still be of questions).
3. I used the mapping to map everyone's favorite session's name to a concise category of program.
4. I join the member background table with the table of favorite program of each member and run R decision tree algorithm.

### Intepretation and conclusion: 
Since this classification task has no obvious unequal consequences depending on false positives/negatives, I interpret these cases equally. The tree is outperforming random guessing / guessing the most frequent option. However, it is still not quite accurate; this suggests that there are 3 possibilities, 1. the variables I fed are not good indicators of program type favor 2. the model is too simplematic to capture the relationship 3. These demographic variables have a lot of missing values, which make it hard to apply to the population.

### Data expectation:
1. less missing values
2. 

```{r}
library(tidyverse)
library(clock)
library(rpart)
library(partykit)
setwd("~/sta130_project/")

```

```{r}
service_table <- read_csv("Service_Deliveries.csv")
table(service_table$service)
# find too many programs, some categories are too specific
```
```{r}
data_patient_members <- read_csv("v3_wrangled_Member_Background.csv")

## Parse members data
## Types of cancer - convert to set of indicator variables for different types of cancer (don't include ones which are too rare)
data_patient_members <- data_patient_members |>
   mutate(lung = str_detect(types_of_cancer, "Lung"),
         breast = str_detect(types_of_cancer, "Breast"),
         brain = str_detect(types_of_cancer, "Brain"),
         leukemia = str_detect(types_of_cancer, "Leukemia"),
         prostate = str_detect(types_of_cancer, "Prostate"),
         colorectal = str_detect(types_of_cancer, "Colorectal"),
         pancreatic = str_detect(types_of_cancer, "Pancreatic"),
         lymphoma_non_hodgkins = str_detect(types_of_cancer, "Lymphoma, Non-hodgkins"),
         liver = str_detect(types_of_cancer, "Liver"),
         multiple_myeloma = str_detect(types_of_cancer, "Multiple Myeloma"),
         bladder = str_detect(types_of_cancer, "Bladder"),
         cervical = str_detect(types_of_cancer, "Cervical"),
         esophagus = str_detect(types_of_cancer, "Esophagus"),
         gallbladder_and_bile_duct = str_detect(types_of_cancer, "Gallbladder and Bile Duct"),
         head_and_neck = str_detect(types_of_cancer, "Head and Neck"),
         kidney = str_detect(types_of_cancer, "Kidney"),
         lymphoma_hodgkins = str_detect(types_of_cancer, "Lymphoma, Hodgkins"),
         ovarian = str_detect(types_of_cancer, "Ovarian"),
         skin = str_detect(types_of_cancer, "Skin"),
         stomach = str_detect(types_of_cancer, "Stomach"),
         testicular = str_detect(types_of_cancer, "Testicular"),
         thyroid = str_detect(types_of_cancer, "Thyroid"),
         cervical = str_detect(types_of_cancer, "Cervical"),
         other = str_detect(types_of_cancer, "Other"))


## Impairments - convert to set of indicator variables for different types of impairment (don't include ones which are too rare)
data_patient_members <- data_patient_members |>
  mutate(visual = str_detect(i_have_the_following_impairment, "visual"),
         cognitive = str_detect(i_have_the_following_impairment, "cognitive"),
         limited_mobility = str_detect(i_have_the_following_impairment, "mobility"),
         none_of_the_above = str_detect(i_have_the_following_impairment, "None of the above"),
         other = str_detect(i_have_the_following_impairment, "other"),
         speech = str_detect(i_have_the_following_impairment, "speech"),
         hearing = str_detect(i_have_the_following_impairment, "hearing")) 

## Program interests: Split into boolean indicators for different categories
data_patient_members <- data_patient_members |>
  mutate(therapeutic_arts = str_detect(program_interest, "Therapeutic Arts"),
         educational_and_self_developoment = str_detect(program_interest, "Educational & Self-Development"),
         symptom_management = str_detect(program_interest, "Symptom Management"),
         social_events = str_detect(program_interest, "Social / Events"),
         financiaL_and_workplace_strategies = str_detect(program_interest, "Financial & Workplace Strategies"),
         exercise_and_movement = str_detect(program_interest, "Exercise & Movement"),
         individual_and_group_support = str_detect(program_interest, "Individual & Group Support")) |>
  mutate(num_program_interests = therapeutic_arts + educational_and_self_developoment + symptom_management + social_events + financiaL_and_workplace_strategies + exercise_and_movement + individual_and_group_support)

## How did you hear about wellspring: Split into booleans
data_patient_members <- data_patient_members |>
  mutate(word_of_mouth = str_detect(how_have_you_heard_of_wellspring, "Word of mouth"),
         flyer = str_detect(how_have_you_heard_of_wellspring, "flyer"),
         wellspring_member = str_detect(how_have_you_heard_of_wellspring, "Member"),
         wellspring_member = str_detect(how_have_you_heard_of_wellspring, "Volunteer"),
         doctor = str_detect(how_have_you_heard_of_wellspring, "Doctor"),
         hospital = str_detect(how_have_you_heard_of_wellspring, "hospital"),
         hospital = str_detect(how_have_you_heard_of_wellspring, "healthcare"),
         tv = str_detect(how_have_you_heard_of_wellspring, "television"),
         website = str_detect(how_have_you_heard_of_wellspring, "Website"),
         family = str_detect(how_have_you_heard_of_wellspring, "Family"),
         social_media = str_detect(how_have_you_heard_of_wellspring, "Social Media"),
         other = str_detect(how_have_you_heard_of_wellspring, "Other"))
```

```{r}
service_table_integrated <- service_table |> mutate(organized_service = sapply(strsplit(service, ":"), `[`, 1))
service_table_integrated |> group_by(member_id) |> 
  summarise(fav_program = first(names(sort(table(organized_service), decreasing = TRUE)))) |>
  group_by(fav_program) |> summarise(n = n())
# the output is given to ai to generate the mapping in next part
# to classify programs that are fav_program of some members
service_table_integrated |> group_by(member_id, organized_service) |> 
  summarise(n = n())
```

```{r}
program_category_mapping_concise <- data.frame(
  organized_service = c(
    "Art Therapy", "Art Workshop", "Art Workshops", "Be Well Talks", "Bereavement Support Group - Family",
    "Bereavement Support Group - Spousal/Partner", "Bladder Cancer Peer Support", "Brain Fog",
    "Breast Cancer Support Group", "Breathing Wellness", "Cancer Coaching", "Cancer Coaching Volunteer Training",
    "Cancer Exercise", "Cancer Related Fatigue", "Cancer Transition Coaching", "Caregivers Connect",
    "Chair Yoga", "Chi Life", "Child Minding", "Children's and Parents' Program",
    "Dance and Flow Workshop", "Drumming Workshop", "ELLICSR - Healing Journey 1", "ELLICSR - Yoga",
    "Employment Law & LTD Legal Clinic", "Energy Sessions (Reiki + Healing Touch + Therapeutic Touch)",
    "Energy Sessions (Reiki)", "Family Counselling", "Family Event", "Family Navigation",
    "Gay Men's Cancer Support Group", "Gay and Queer Men's Cancer Support Group",
    "Groupe de soutien - Cancer de l’ovaire", "Groupe de soutien du cancer de l'ovaire", "Healing Journey 1",
    "Healing Journey 2", "Healing Journey 3", "Healing Journey 5", "Healing Journey 6",
    "Healing Journey Level 1", "Healing Journey Level 2", "Healing Journey Level 3",
    "Lesbians and Queer Support Group", "Let's Knit", "Let's Sing", "Living With Chronic Cancer Support Group",
    "Look Good Feel Better", "Look Good Feel Better (external)", "Lymphoma Support Group", "Meditation",
    "Meditation Workshop", "Meditation Workshops", "Members' Potluck", "Mindfulness Based Cancer Recovery",
    "Money Matters", "Money Matters (Ontario)", "Money Matters Professional Consultations", "Music Therapy",
    "Music Workshop", "Myeloma Peer Support", "Nourish", "OASIS Parents Support Group",
    "Oasis Advanced Cancer Support Group", "Oasis Art Expression Studio", "Ovarian Cancer Circle",
    "Ovarian Cancer Peer Support", "Ovarian Cancer Peer Support - French", "Ovarian Cancer Support Group",
    "Ovarian Cancer Support Group - French", "Pancreatic Cancer Peer Support",
    "Pancreatic Cancer Short Term Counselling", "Pancreatic Cancer Short-term Professional Counselling",
    "Pancreatic Cancer Support Group", "Patient Support Circle", "Patient's Connect", "Peer Support",
    "Peer Support - L'Chaim", "Peer Support - Ovarian Cancer", "Peer Support - Pancreatic Cancer",
    "Peer Support - Welcome Calls", "Prostate Cancer Support Group (external)",
    "Prostate Cancer Support Group - Oakville/Mississauga", "Reiki", "Relaxation and Visualization",
    "Returning to Work", "Short Term Counselling", "Short-term Professional Counselling",
    "Sound Bath Meditations Workshop", "St. Michael's - Healing Journey 1", "Support Circle",
    "Support Group", "Volunteer Recognition Event", "Volunteer Support Training",
    "Volunteer Support Training (Fee For Service)", "Wellness Workshop", "Wellness Workshops",
    "Wellness at Work", "Wellspring Discussion Forum", "Wellspring Exercise 101", "Wig Program",
    "Writing Workshop", "Writing Workshops", "Yoga"
  ),
  program_category = c(
    "Creative & Expressive Arts", "Creative & Expressive Arts", "Creative & Expressive Arts",
    "Wellness Education & Skills", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Mind-Body Wellness", "Peer & Community Support", "Wellness Education & Skills",
    "Mind-Body Wellness", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Mind-Body Wellness", "Creative & Expressive Arts",
    "Family & Caregiver Resources", "Family & Caregiver Resources", "Creative & Expressive Arts",
    "Creative & Expressive Arts", "Peer & Community Support", "Mind-Body Wellness",
    "Wellness Education & Skills", "Mind-Body Wellness", "Mind-Body Wellness",
    "Family & Caregiver Resources", "Information & Resource Events", "Family & Caregiver Resources",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Creative & Expressive Arts", "Creative & Expressive Arts",
    "Peer & Community Support", "Wellness Education & Skills", "Wellness Education & Skills",
    "Peer & Community Support", "Mind-Body Wellness", "Mind-Body Wellness",
    "Mind-Body Wellness", "Information & Resource Events", "Mind-Body Wellness",
    "Wellness Education & Skills", "Wellness Education & Skills", "Wellness Education & Skills",
    "Creative & Expressive Arts", "Creative & Expressive Arts", "Peer & Community Support",
    "Peer & Community Support", "Family & Caregiver Resources", "Peer & Community Support",
    "Creative & Expressive Arts", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Mind-Body Wellness", "Mind-Body Wellness",
    "Wellness Education & Skills", "Wellness Education & Skills", "Wellness Education & Skills",
    "Mind-Body Wellness", "Peer & Community Support", "Peer & Community Support",
    "Peer & Community Support", "Information & Resource Events", "Wellness Education & Skills",
    "Wellness Education & Skills", "Wellness Education & Skills", "Wellness Education & Skills",
    "Wellness Education & Skills", "Peer & Community Support", "Wellness Education & Skills",
    "Information & Resource Events", "Creative & Expressive Arts", "Creative & Expressive Arts",
    "Mind-Body Wellness"
  )
)


# Display the corrected and final data frame
print(program_category_mapping_concise)
```

```{r}
service_table_added_category <- service_table_integrated |> inner_join(program_category_mapping_concise)
summary_table_mem_fav_program <- service_table_added_category |> group_by(member_id) |>
  summarise(fav_program_category = first(names(sort(table(program_category), decreasing = TRUE))))
summary_table_mem_fav_program |> glimpse()
```

```{r}
mem_bg_data <- read_csv("v4_wrangled_Member_Background.csv")
final_data_1 <- mem_bg_data |> right_join(summary_table_mem_fav_program)
final_data_2 <- mem_bg_data |> inner_join(summary_table_mem_fav_program)

final_data_1 |> anti_join(final_data_2) # choose final_data_2 since the result of this are no profile members

final_data <- final_data_2 |> dplyr::select(-member_id, -mailing_city, -member_start_year, -member_start_month, -i_have_the_following_impairment, -types_of_cancer, -how_have_you_heard_of_wellspring, -number_of_present_service_deliveries, -number_of_absent_service_deliveries, -last_service_date_year, -last_service_date_month, -last_service_date_day, -program_interest, -hospitals_where_you_were_treated)
```

```{r}
# traning on entire dataset
model <- rpart(fav_program_category ~ , data=final_data)
predictions <- predict(model, newdata=final_data_2, type="class")
table(predictions, final_data_2$fav_program_category)
plot(as.party(model), gp=gpar(cex=0.6), type = "simple")


n_obs = nrow(final_data_2)
summ <- final_data_2 |> group_by(fav_program_category) |>
  summarize(fav_program_category_prop = n()/n_obs)

summ |> ggplot(aes(x=fav_program_category, y=fav_program_category_prop)) + 
  geom_col() + labs(x="fav program type", y="propotion") + coord_flip()

plot_helper <- final_data_2 |> group_by(gender, fav_program_category) |>
  summarize(fav_program_category_prop = n()/n_obs)
plot_helper |> ggplot(aes(gender, fav_program_category_prop, fill = fav_program_category)) + geom_col() +
  labs(x="gender", y="proportion within all observations") + coord_flip()

plot_helper <- final_data_2 |> filter(gender=="Unknown") |> group_by(mailing_state_province, fav_program_category) |>
  summarize(fav_program_category_prop = n()/n_obs)
plot_helper |> ggplot(aes(mailing_state_province, fav_program_category_prop, fill = fav_program_category)) + geom_col() +
  labs(x="residencial state / province", y="proportion within all observations") + coord_flip()

accuracy_of_guessing_PC_1 <- max(summ$fav_program_category_prop)
accuracy_of_model_1 <- 2290 / n_obs
accuracy_of_model_1 > accuracy_of_guessing_PC_1

```

```{r}
# traning on people who are with disabilities
library(rpart.plot)
model <- rpart(fav_program_category ~ gender + age_years + parent_of_a_child_under_18 + i_identify_as_lgbtq + i_identify_as_poc + mailing_state_province, data=final_data)
predictions <- predict(model, newdata=final_data, type="class")
table(predictions, final_data$fav_program_category)

plot(as.party(model),  gp=gpar(cex=0.4), type="s", main="Decision Tree")
rpart.plot(model)

n_obs = nrow(final_data)
summ <- final_data |> group_by(fav_program_category) |>
  summarize(fav_program_category_prop = n()/n_obs)

summ |> ggplot(aes(x=fav_program_category, y=fav_program_category_prop)) + 
  geom_col() + labs(x="fav program type", y="propotion") + coord_flip()

plot_helper <- final_data |> group_by(gender, fav_program_category) |>
  summarize(fav_program_category_prop = n()/n_obs)
plot_helper |> ggplot(aes(gender, fav_program_category_prop, fill = fav_program_category)) + geom_col() +
  labs(x="gender", y="proportion within all observations") + coord_flip()

accuracy_of_guessing_PC_2 <- max(summ$fav_program_category_prop)
accuracy_of_model_2 <- (17 + 42) / n_obs
accuracy_of_model_2 > accuracy_of_guessing_PC_2

plot_helper <- final_data |> filter(gender=="Unknown") |> group_by(mailing_state_province, fav_program_category) |>
  summarize(fav_program_category_prop = n()/n_obs)
plot_helper |> ggplot(aes(mailing_state_province, fav_program_category_prop, fill = fav_program_category)) + geom_col() +
  labs(x="residencial state / province", y="proportion within all observations") + coord_flip()


```
